Tests of measurement invariance without subgroups: A~generalization of classical methods

### 2012-05-31 www/paper.Rnw
Improved paper.Rnw so that instead of the default red boundaries, gray boundaries are plotted. Requires strucchange >= 1.4-7.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@130 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<<preliminaries>>=
library("OpenMx")
library("lavaan")
library("strucchange")
library("mvtnorm")
library("lattice")
source("mz.R")
source("lavaan-S3.R")
source("sim.R")
@

@@

\begin{figure}
\caption{A comparison of the $\max \mathit{LM}$ (solid line) and $\max \mathit{LR}$ (dashed line) test
  statistics for~\eqref{eq:h0} (i.e., $k^* = 19$).  The solid horizontal line
  corresponds to the critical value at $\alpha= 0.05$ while the dotted vertical
  line highlights the threshold at which both test statistics assume their maximum.}
\label{fig:lrlm}
\setkeys{Gin}{width=\textwidth}
<<lm-lr-19, fig=TRUE, height=5, width=6>>=
plot(gefp_19_info, functional = supLM(0.1),
  xlab = "Age", ylab = "LR and LM statistics (k* = 19)", main = "", ylim = c(0, 50))
legend("topleft", c("LR", "LM"), lty = c(2, 1), bty = "n")
lines(mz.age, mz.lrstat, lty = 2)
abline(v = mz.age[which.max(mz.lrstat)], lty = 3)
@
\end{figure}

@@

\begin{figure}
\centering
\caption{Cumulative score processes for each verbal factor loading. 
  The horizontal lines correspond to the critical value of the
  double-max test at $\alpha= 0.05$.}
\label{fig:lambdas}
\setkeys{Gin}{width=0.8\textwidth}
<<gefp-3-visualization, fig=TRUE, height=8, width=5.5>>=
plot(gefp_3_info,  functional = maxBB, aggregate = FALSE, ylim = c(-2, 2),
  main = "DM, k* = 3", xlab = "Age", ylab = expression(lambda[11], lambda[21], lambda[31]))
@
\end{figure}

@@

\begin{figure}
\caption{Three test statistics of~\eqref{eq:simh0} (with $k^* = 3$)
  and~\eqref{eq:h0} (with $k^* = 19$), 
  based on the example involving measurement invariance with respect
  to student age.  Solid, horizontal lines represent critical values
  at $\alpha = 0.05$, and the dotted, horizontal lines (second row)
  represent values of the Cram\'{e}r-von Mises test statistic.}
\label{fig:cusumex}
\setkeys{Gin}{width=\textwidth}
<<gefp-3-19-visualization, fig=TRUE, height=8, width=6>>=
par(mfcol = c(3, 2))
plot(gefp_3_info,  functional = maxBB, main = "DM, k* = 3", xlab = "Age")
plot(gefp_3_info,  functional = meanL2BB, main = "CvM, k* = 3", xlab = "Age")
plot(gefp_3_info,  functional = supLM(0.1), main = "max LM, k* = 3", xlab = "Age")
plot(gefp_19_info, functional = maxBB, main = "DM, k* = 19", xlab = "Age")
plot(gefp_19_info, functional = meanL2BB, main = "CvM, k* = 19", xlab = "Age")
plot(gefp_19_info, functional = supLM(0.1), main = "max LM, k* = 19", xlab = "Age")
@
\end{figure}

@@

\begin{figure}
\centering
\caption{Cumulative score processes for the four group-specific parameters across GPA
  in the ``minority, stereotype prime'' group
  using data from Study 1 of Wicherts et al.\ (2005). %% \citeA{WicDol05} does not work in caption
  The horizontal lines correspond to the critical value of
  the doule-max test at $\alpha = 0.05$.}
\label{fig:cusumgpa}
\setkeys{Gin}{width=0.8\textwidth}
<<gefp-4-plot, fig=TRUE, height=8, width=5.5>>=
plot(gefp_4_info,  functional = maxBB, aggregate = FALSE,
  ylim = c(-2.15, 2.15), main = "DM, k* = 4", xlab = "GPA",
  ylab = expression(lambda[num], eta, mu[num], psi[num]))
@
\end{figure}

>>>>

<<preliminaries>>=
## packages
library("OpenMx")
library("lavaan")
library("strucchange")
library("mvtnorm")
library("lattice")

## auxiliary code
source("mz.R")
source("estfun-lavaan.R")
source("sim.R")

## convenience function for plotting boundaries without color
get_boundary <- function(obj, fun, trim = 0) {
  bound <- fun$boundary(0:obj$nobs/obj$nobs)
  bound <- fun$computeCritval(0.05, ncol(obj$process)) * bound
  bound <- zoo(bound, time(obj$process))
  if(trim > 0) bound <- head(tail(bound, - floor(trim * obj$nobs)), - floor(trim * obj$nobs))
  return(bound)
}
@

@@

\begin{figure}
\caption{A comparison of the $\max \mathit{LM}$ (solid line) and $\max \mathit{LR}$ (dashed line) test
  statistics for~\eqref{eq:h0} (i.e., $k^* = 19$).  The gray horizontal line
  corresponds to the critical value at $\alpha= 0.05$ while the dotted vertical
  line highlights the threshold at which both test statistics assume their maximum.}
\label{fig:lrlm}
\setkeys{Gin}{width=\textwidth}
<<lm-lr-19, fig=TRUE, height=5, width=6>>=
plot(gefp_19_info, functional = supLM(0.1),
  xlab = "Age", ylab = "LR and LM statistics (k* = 19)",
  main = "", ylim = c(0, 50), boundary = FALSE)
legend("topleft", c("LR", "LM"), lty = c(2, 1), bty = "n")
lines(mz.age, mz.lrstat, lty = 2)
lines(get_boundary(gefp_19_info, supLM(0.1), 0.1), col = "slategray", lwd = 1.7)
abline(v = mz.age[which.max(mz.lrstat)], lty = 3)
@
\end{figure}

@@

\begin{figure}
\centering
\caption{Cumulative score processes for each verbal factor loading. 
  The solid gray horizontal lines correspond to the critical value of the
  double-max test at $\alpha= 0.05$.}
\label{fig:lambdas}
\setkeys{Gin}{width=0.8\textwidth}
<<gefp-3-visualization, fig=TRUE, height=8, width=5.5>>=
gefp_3_bound <- get_boundary(gefp_3_info, maxBB)
plot(gefp_3_info$process,  ylim = c(-2, 2),
  main = "DM, k* = 3", xlab = "Age", ylab = expression(lambda[11], lambda[21], lambda[31]),
  panel = function(x, y, ...) {
    abline(h = 0)
    lines(gefp_3_bound, col = "slategray", lwd = 1.7)
    lines(-gefp_3_bound, col = "slategray", lwd = 1.7)
    lines(x, y, ...)
  })
@
\end{figure}

@@

\begin{figure}
\caption{Three test statistics of~\eqref{eq:simh0} (with $k^* = 3$)
  and~\eqref{eq:h0} (with $k^* = 19$), 
  based on the example involving measurement invariance with respect
  to student age.  Solid gray horizontal lines represent critical values
  at $\alpha = 0.05$, and the dotted, horizontal lines (second row)
  represent values of the Cram\'{e}r-von Mises test statistic.}
\label{fig:cusumex}
\setkeys{Gin}{width=\textwidth}
<<gefp-3-19-visualization, fig=TRUE, height=8, width=6>>=
par(mfcol = c(3, 2))
plot(gefp_3_info,  functional = maxBB, main = "DM, k* = 3", xlab = "Age", boundary = FALSE)
lines(get_boundary(gefp_3_info, maxBB), col = "slategray", lwd = 1.7)
plot(gefp_3_info,  functional = meanL2BB, main = "CvM, k* = 3", xlab = "Age", boundary = FALSE)
lines(get_boundary(gefp_3_info, meanL2BB), col = "slategray", lwd = 1.7)
plot(gefp_3_info,  functional = supLM(0.1), main = "max LM, k* = 3", xlab = "Age", boundary = FALSE)
lines(get_boundary(gefp_3_info, supLM(0.1), 0.1), col = "slategray", lwd = 1.7)
plot(gefp_19_info, functional = maxBB, main = "DM, k* = 19", xlab = "Age", boundary = FALSE)
lines(get_boundary(gefp_19_info, maxBB), col = "slategray", lwd = 1.7)
plot(gefp_19_info, functional = meanL2BB, main = "CvM, k* = 19", xlab = "Age", boundary = FALSE)
lines(get_boundary(gefp_19_info, meanL2BB), col = "slategray", lwd = 1.7)
plot(gefp_19_info, functional = supLM(0.1), main = "max LM, k* = 19", xlab = "Age", boundary = FALSE)
lines(get_boundary(gefp_19_info, supLM(0.1), 0.1), col = "slategray", lwd = 1.7)
@
\end{figure}

@@

\begin{figure}
\centering
\caption{Cumulative score processes for the four group-specific parameters across GPA
  in the ``minority, stereotype prime'' group
  using data from Study 1 of Wicherts et al.\ (2005). %% \citeA{WicDol05} does not work in caption
  The solid gray horizontal lines correspond to the critical value of
  the doule-max test at $\alpha = 0.05$.}
\label{fig:cusumgpa}
\setkeys{Gin}{width=0.8\textwidth}
<<gefp-4-plot, fig=TRUE, height=8, width=5.5>>=
gefp_4_bound <- get_boundary(gefp_4_info, maxBB)
plot(gefp_4_info$process,
  ylim = c(-2.15, 2.15), main = "DM, k* = 4", xlab = "GPA",
  ylab = expression(lambda[num], eta, mu[num], psi[num]),
  panel = function(x, y, ...) {
    abline(h = 0)
    lines(gefp_4_bound, col = "slategray", lwd = 1.7)
    lines(-gefp_4_bound, col = "slategray", lwd = 1.7)
    lines(x, y, ...)
  })

@
\end{figure}
